Video-Agnostic Perturbations: Efficient Targeted Attacks \\ for Siamese Visual Tracking

### 2021-01-15 博士/论文提交/adversarial_attack/IJCAI/FormattingGuidelines-IJCAI-21/main.tex
高晋老师 20210115 修改 introduction

Object tracking refers to the task of sequentially locating a specified moving object in a video, given only its initial state. Recently, Siamese networks \cite{ATOM} \cite{SiamFC} have demonstrated a significant improvement in object tracking performances. Siamese trackers formulate the visual object tracking problem as learning cross-correlation similarities between a target template and a search region. Tracking is then performed by finding the target object from the search image region by computing the highest visual similarity.

Recently, the robustness of Siamese trackers have recently attracted a lot of attention \cite{TemporallyTransferablePerturbations,efficient,SPARK}. Despite the impressive performance of Siamese architectures on challenging visual object tracking task, these trackers were shown to be highly vulnerable to perturbations.
\cite{TemporallyTransferablePerturbations} propose a framework to generate a single temporally transferable adversarial perturbation from the object template image only. This perturbation can then be added to every search image to adversarially attack the subsequent video frames.
\cite{efficient} present an end-to-end network FAN (Fast Attack Network) that uses a novel drift loss combined with the embedded feature loss to attack the Siamese network based trackers.
\cite{SPARK} propose the spatial-aware online incremental attack (a.k.a. SPARK) that performs spatial-temporal sparse incremental perturbations online and makes the adversarial attack less perceptible.

All the above methods can fool object trackers successfully. However, these attack methods are not suitable for the real-world online-tracking phase where we can not get access to the limited computational resources.
The reason is, visual object tracking is a computationally intensive task which usually needs to be run in real time, but the computational resources of small platforms (e.g., mobile, robotic, self-driving car) used to deploy the trackers are often limited.
Therefore, the attack approaches for object tracking should occupy as little computing resources as possible.
But the reality is that most attack algorithms either perform iterative optimisation to calculate adversarial examples, or generate perturbations through Generative Adversarial Networks which occupy the GPU memory.

Universal adversarial perturbation (UAP) \cite{UAP} is an image-agnostic perturbation which can fool most images from a data distribution. Being universal, UAPs can be conveniently exploited to perturb unseen datapoints on-the-fly without extra computation. Therefore, UAPs are particularly useful when attacking real-time applications deployed on platforms with limited computing resources.
However, no existing work has touched the topic of attacking the Siamese trackers using UAPs, because it is difficult to apply existing UAPs in Siamese trackers directly. The reasons come from two aspects.
\begin{itemize}
\item Different input. Most UAPs are designed for typical neural networks, which usually accept one image as input; while Siamese-based tracking networks accept two inputs: the template image and the search image.
\item Different goals. The goal of existing UAP methods is to disturb unary and binary model outputs for single instance; while we need to use universal perturbations to mislead Siamese trackers along with a specified trajectory.
\end{itemize}

In this paper, we make the first attempt in finding the universal perturbations that fool a state-of-the-art deep Siamese trackers on all natural videos. We show in this paper the existence of such universal perturbations that cause the tracker to track object along any specified trajectory.
Specifically, we aim to attack the tracker by adding an imperceptible perturbation to the template image and pasting a \textit{fake target}, i.e., a small adversarial patch into the search image adhering to the predefined trajectory, so that tracker outputs the location/size of the fake target instead of the real target.

Different from the existing attack methods which are intrinsic dependent on videos, we generate the perturbations that can fool the network on most natural videos. Perturbing a new target then only involves the mere addition/paste operation of the universal perturbations to the template/search image -- and does not require solving an optimisation problem or network inference.
Overall, our contributions can be summarized as follows:
\begin{itemize}
\item We show the existence of universal video-agnostic perturbations for state-of-the-art deep Siamese trackers. To the best of our knowledge, this is the first work to perform universal targeted attack against the visual object tracking task.
\item We propose an end-to-end pipeline to generate the universal perturbations using large scale object tracking dataset.
\item Unlike most existing attack methods which only attack the classification branch of the tracker, we simultaneously manipulate the classification and bounding box regression result of the Siamese tracker.
\item Experiment results on OTB2015 \cite{OTB}, GOT-10k \cite{GOT-10k} and LaSOT \cite{GOT-10k} datasets demonstrate that our method can effectively fool the Siamese trackers in a targeted attack manner, which comes at virtually no cost.
\end{itemize}

>>>>

Given an arbitrary detected or annotated object of interest in the initial video frame, visual object tracking is aimed at {\it recognizing} and {\it localizing} other instances of the same object in subsequent frames. This paradigm of tracking visual objects from a single initial exemplar in the testing phase has been broadly cast as a Siamese network-based one-shot problem recently \cite{SiamFC,SiamRPN,DaSiamRPN,SiamRPN++,SiamMask,SiamFC++}, which is termed as Siamese visual tracking and recognised as highly effective and efficient for visual tracking.

Recently, the robustness of Siamese trackers have attracted much attention in the sense of testing their vulnerability to adversarial attacks, and the focus has been directed towards more efficient and low-cost attacks \cite{TemporallyTransferablePerturbations,efficient,SPARK}. Despite their success, these attack methods are not suitable for attacking the real-world online-tracking systems built upon the small edge platforms with limited computational resources. The reason is that they still need to craft the perturbations for each video independently based on either iterative optimisation or adversarial network inference, for which the adequacy of computational resources may be not ensured by the computationally intensive tracking systems.

Universal adversarial perturbations (UAPs) proposed in \cite{UAP} can fool most images from a data distribution in an image-agnostic manner. Being universal, UAPs can be conveniently exploited to perturb unseen data on-the-fly without extra computations. Therefore, UAPs are particularly useful when attacking real-time applications deployed on platforms with limited computational resources. However, no existing work has touched the topic of attacking the Siamese trackers using UAPs, because it is hard to apply existing UAPs to attack Siamese trackers directly. The main reason lies in the fact that, (a) most UAPs are designed for typical neural networks with one image as input while Siamese networks accept both the template and search images, and (b) the goal of existing UAP methods is to disturb unary and binary model outputs for single instance while we need to use universal perturbations to mislead Siamese trackers to follow a specified trajectory.

In this paper, we make the first attempt in finding the video-agnostic perturbations that fool state-of-the-art Siamese trackers in a targeted attack manner, which comes at virtually no cost in the online-tracking phase. Specifically, we aim to attack the trackers by adding an universal imperceptible perturbation to the template image and pasting a \textit{fake target}, \ie, a small universal adversarial patch, into the search image adhering to the predefined trajectory, so that tracker outputs the location and size of the \textit{fake target} instead of the real target. Our generated video-agnostic perturbations allow perturbing a novel video to come at no additional cost except the mere addition and pasting operations -- and not require gradient optimization or network inference. Experiment results on OTB2015 \cite{OTB}, GOT-10k \cite{GOT-10k} and LaSOT \cite{GOT-10k} datasets demonstrate the effectiveness and efficiency of our approach.
